---
title: "STA 325 - Data Expedition"
author: "Ethan Shen, Austin Jia, Malavi Ravindran, Steven Herrera"
date: "10/26/2019"
output: rmdformats::readthedown
---

```{r, echo = FALSE, include = FALSE}
# Installing packages
pkgTest <- function(x) {
  if (!require(x,character.only = TRUE)) {
    install.packages(x,dep=TRUE)
  }
}
pkgs <- c("ggplot2", "cowplot", "LaplacesDemon", "gam", "splines", "mgcv", "polywog", "DAAG")
for (pkg in pkgs) {
  pkgTest(pkg)
}
```

```{r, include = FALSE, echo = FALSE}
library(ggplot2)
library(cowplot)
library(LaplacesDemon)
library(gam)
library(splines)
library(mgcv)
library(polywog)
library(DAAG)
```

```{r}
train <- read.csv("data-train.csv")
test <- read.csv("data-test.csv")
```

# Exploratory Data Analysis

```{r}
train$Fr <- invlogit(train$Fr)

p1 <- qplot(train$St, bins = 20, 
      main = "Distribution of St", 
      xlab = "St", 
      ylab = "Frequency")
p2 <- qplot(train$Re, bins = 20,
      main = "Distribution of Re", 
      xlab = "Re", 
      ylab = "Frequency")
p3 <- qplot(train$Fr, bins = 20,
      main = "Distribution of Fr", 
      xlab = "Fr", 
      ylab = "Frequency")
cowplot::plot_grid(p1,p2,p3)
```

# Moment 1 - Steven

### General Framework

### Linear Model (log transformations)


#### Using Ridge and LASSO regression 

### Polynomial 

### GAMs

### Splines

### Random Forest

### MSE Scores LATEX Table

### Flexibility vs. MSE Table

### Final Insights



# Moment 2 - Malavi

### General Framework

### Linear Model (log transformations)


#### Using Ridge and LASSO regression 

### Polynomial 

### GAMs

### Splines

### Random Forest

### Cross Validation

### MSE Scores LATEX Table

### Flexibility vs. MSE Table

### Final Insights



# Moment 3 - Austin

### General Framework

### Linear Model (log transformations)

```{r}
lm.fit.3 <- lm(R_moment_3 ~ St + Re + Fr, data = train) 
summary(lm.fit.3)
predict(lm.fit.3, list(St = test$St, Re = test$Re, Fr = test$Fr))
plot(lm.fit.3)


cv.lm(train, form.lm = formula(log(R_moment_3) ~ St + Re + Fr), m = 10)
```

### Polynomial 

```{r}
cv1 <- cv.polywog(R_moment_3 ~ St + Re + Fr,
                  data = train,
                  degrees.cv = 1:4,
                  nfolds = 5,
                  thresh = 1e-4)
print(cv1)

## Extract best model and bootstrap
fit1 <- cv1$polywog.fit
fit1 <- bootPolywog(fit1, nboot = 5)
summary(fit1)
```


### GAMs

```{r}
qplot(R_moment_3, data = train,
      main = "Distribution of Moment 3", 
      xlab = "Moment 3", 
      ylab = "Frequency")

qplot(log(R_moment_3), data = train,
      main = "Distribution of log(Moment 3)", 
      xlab = "log(Moment 3)", 
      ylab = "Frequency")

```

```{r, include=FALSE}
cross_validate <- function(model_type, formula, data, k, log_response) {
  #Randomly shuffle the data
  shuffled_data<-data[sample(nrow(data)),]
  #Create 10 equally size folds
  folds <- cut(seq(1,nrow(shuffled_data)),breaks=k,labels=FALSE)
  #Perform 10 fold cross validation
  avg_mse <- 0
  for(i in 1:k){
    #Segement your data by fold using the which() function 
    testIndexes <- which(folds==i,arr.ind=TRUE)
    testData <- shuffled_data[testIndexes, ]
    trainData <- shuffled_data[-testIndexes, ]
    
    model <- model_type(as.formula(formula), data=trainData)
    # validate model
    if (log_response) {
      test_mse <-  mean((exp(predict(model, testData)) - testData$BWTG_C)^2)
    }
    else {
      test_mse <- mean((predict(model, testData) - testData$BWTG_C)^2)
    }
    avg_mse <- avg_mse + test_mse
    }
  print('Average test MSE')
  print(avg_mse / k)
  return (avg_mse / k)
}
```

```{r}
gam.1.3 <- gam(log(R_moment_3) ~ s(St) + Re + Fr, data = train)
summary(gam.1.3)
gam.2.3 <- gam(log(R_moment_3) ~ s(log(St)) + Re + Fr, data = train)
summary(gam.2.3) 
```

Logging `St` definitely helps

```{r}
#GOOD MODEL 
gam.3.3 <- gam(log(R_moment_3) ~ s(log(St)) + Re + Fr + Re * Fr, data = train)
summary(gam.3.3)
anova(gam.2.3, gam.3.3, test = "F")
plot(gam.3.3, pages=1, residuals=TRUE, all.terms=TRUE, shade=TRUE, shade.col=2)
```


### Splines

```{r}
spline.1 <- lm(R_moment_3 ~ ns(St, df = 2) + ns(Re, df = 4) + ns(Fr, df=2), data = train) #.3942
spline.2 <- lm(R_moment_3 ~ ns(log(St), df = 4) + Re + ns(Fr, df=2), data = train) #.3366
spline.3 <- lm(log(R_moment_3) ~ ns(log(St) + Re, df = 4) + ns(Fr, df=2), data = train) #.4697, AIC 4410.8
spline.4 <- lm(R_moment_3 ~ bs(St + Re, df = 5) + bs(Fr, degree = 3), data = train) #.4607

summary(spline.1)
summary(spline.2)
summary(spline.3)
summary(spline.4)
```

### Random Forest

### MSE Scores LATEX Table

### Flexibility vs. MSE Table

### Final Insights




# Moment 4 - Ethan 

### General Framework

### Linear Model (log transformations)

```{r}
lm.fit.4 <- lm(log(R_moment_4) ~ St + Re + Fr, data = train) 
summary(lm.fit.4)
library(DAAG)
cv.lm(train, form.lm = formula(log(R_moment_4) ~ St + Re + Fr), m = 5)
```

#### Using Ridge and LASSO regression 

### Polynomial 

```{r}
unique(train$St)
poly.fit.4 <- lm(R_moment_4 ~ poly(St,2) + poly(Re,2) + poly(Fr, 2), data = train) 
summary(poly.fit.4)
library(polywog)

cv.polywog(R_moment_3 ~ St + Re + Fr,
           data = train,
           degrees.cv = 1:4,
           nfolds = 5,
           thresh = 1e-4)
```

### GAMs

```{r}
qplot(R_moment_4, data = train,
      main = "Distribution of Moment 4", 
      xlab = "Moment 4", 
      ylab = "Frequency")

qplot(log(R_moment_4), data = train,
      main = "Distribution of log(Moment 4)", 
      xlab = "log(Moment 4)", 
      ylab = "Frequency")

```

```{r}
gam.1 <- gam(log(R_moment_4) ~ s(St) + Re + Fr, data = train)
summary(gam.1)
gam.2 <- gam(log(R_moment_4) ~ s(log(St)) + Re + Fr, data = train)
summary(gam.2) 
```

Logging `St` definitely helps

```{r}
#GOOD MODEL 
library(mgcv)
gam.3 <- gam(log(R_moment_4) ~ s(log(St)) + Re + Fr + Re * Fr, data = train)
summary(gam.3)
AIC(gam.3)
anova(gam.1, gam.2, gam.3, test = "F")
plot(gam.3, pages=1, residuals=TRUE, all.terms=TRUE, shade=TRUE, shade.col=2)
```
```{r}
gam.check(gam.3)
```

```{r}
library(gamclass)
gamclass::CVgam(log(R_moment_4) ~ s(log(St)) + Re + Fr + Re * Fr, data = train, nfold = 5, method = "GCV.Cp")
```



### Splines

```{r}
spline.3 <- lm(log(R_moment_4) ~ ns(log(St) + Re, df = 4) + ns(Fr, df=2), data = train) #.4697, AIC 4410.8
spline.4 <- lm(log(R_moment_4) ~ ns(St + Re) + bs(Fr), data = train) #.4607

summary(spline.3)
AIC(spline.3)
summary(spline.4)
AIC(spline.4)
anova(spline.3, spline.4, test = "F")
```


### Random Forest

### Cross Validation



### MSE Scores LATEX Table

### Flexibility vs. MSE Table

### Final Insights